Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-05 overview.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

We begin this section by first writing down the complete set of
physical equations that can be solved by \enzo, and then briefly
describe the numerical algorithms that we use to solve these
equations.  This section is intended to be an overview of the
\enzo\ code's capabilities: thus, we gather all of the equations
solved into a single place and provide a brief and high-level
introduction to the numerics of the code.  Detailed descriptions of
the individual components are then provided in
Sections~\ref{sec.amr}-\ref{sec.num.analysis}. In
Table~\ref{tab:variables} we provide a convenient summary of all
variables and physical constants used throughout this manuscript.

@@

The Eulerian equations of ideal magnetohydrodynamics including
gravity, in a coordinate systems comoving with the cosmological
expansion, are given by:

\begin{eqnarray} 
  \frac{\partial \rho}{\partial t} 
  + \frac{1}{a} \nabla \cdot (\rho \vecv) & = & 0
  \label{eq:mass} \\
%
  \frac{\partial \rho \vecv}{\partial t}  
  + \frac{1}{a} \nabla \cdot \left(\rho \vecv \vecv + \myvec{I}p^* - 
    \frac{\vecB \vecB}{a} \right) & = &
  - \frac{\dot{a}}{a} \rho \vecv - \frac{1}{a} \rho \nabla \phi
% 
%          + \frac{1}{a} ( \vecv \cdot \nabla ) \vecv
%    & = & - \frac{\dot{a}}{a} \vecv 
%          - \frac{1}{a \rho} \nabla p
%          - \frac{1}{a \rho} (\nalba \times \vecB) \times \vecB
%          - \frac{1}{a} \nabla \phi, 
  \label{eq:momentum} \\
%
  \frac{\partial E}{\partial t} 
  + \frac{1}{a} \nabla \cdot \left[ (E + p^*) \vecv - 
    \frac{1}{a} \vecB (\vecB \cdot \vecv) \right] & = &
  - \frac{\dot{a}}{a} \left( 2E - \frac{B^2}{2a} \right) - 
  \frac{1}{a} \vecv \cdot \nabla \phi 
  - \Lambda + \Gamma + \frac{1}{a^2} \nabla \cdot \fcond
% 
%         + \frac{1}{a} \vecv \cdot \nabla E
%  & = & - \frac{\dot{a}}{a} (3 \frac{p}{\rho} + {\vecv \cdot \vecv} )
%       - \frac{1}{a \rho } \nabla \cdot (p \vecv) 
%%\nonumber   & &   %uncomment for 2-column
%        - \frac{1}{a} \vecv \cdot \nabla \phi 
%               + \Gamma - \Lambda.
  \label{eq:total_energy}  \\
%
  \frac{\partial \vecB}{\partial t} - 
  \frac{1}{a}  \nabla \times (\vecv \times \vecB) & = & 
  -\frac{\dot{a}}{2 a} \vecB \label{eq:induction}
\end{eqnarray}

%\begin{eqnarray} 
%\frac{\partial \rho}{\partial t} 
%          + \vecv \cdot \nabla \rho
%    & = & - \rho \nabla \cdot \vecv,
%        \label{eq:mass} \\
%%
%\frac{\partial \vecv }{ \partial t} 
%          + ( \vecv \cdot \nabla ) \vecv 
%    & = & - \frac{1}{\rho} \nabla p
%          - \nabla \phi,
%        \label{eq:momentum} \\
%%
% \frac{\partial E}{\partial t} + \vecv \cdot \nabla E
%    & = & - \frac{1}{\rho} \nabla \cdot (p \vecv) 
%          - \vecv \cdot \nabla \phi
%%\nonumber          \\  & &     % uncomment for 2-column
%     + \Gamma - \Lambda
%        \label{eq:total_energy}
%\end{eqnarray}
%
In these equations, $E$, $\rho$, $\vecv$, and $\vecB$ are the comoving
total fluid energy density, comoving baryonic density, peculiar velocity, and
comoving magnetic field strength, respectively.  $\myvec{I}$ is the
identity matrix, and $a$ is the cosmological expansion parameter
(which is discussed in more detail below).  The first equation
represents conservation of mass, the second, conservation of momentum,
and the third, conservation of total fluid (kinetic plus thermal)
energy.  They are respectively, the first, second and third moments of
the Boltzmann equation.  The fourth equation is the magnetic induction
equation.  Terms representing radiative cooling ($\Lambda$) and
heating ($\Gamma$) enter on the right-hand side of the energy equation
(\ref{eq:total_energy}), as well as the flux due to thermal heat
conduction ($\fcond$).

The total fluid energy density $E$ is given by:
\begin{equation}
E =  e + \frac{\rho v^2}{2}  + \frac{B^2}{2a}
        \label{eq:total_energy_def}.
\end{equation}
where $e$ is the comoving thermal energy density, and the total
isotropic pressure $p^*$ is given by
\begin{equation}
p^* = p + \frac{B^2}{2a},
\end{equation}
where $p$ is the gas pressure.  The equations are closed with the
equation of state and Poisson's equation (we work in the Newtonian
limit):
%
\begin{eqnarray}
%
  e &=& p / [(\gamma - 1) \rho ],
  \label{eq:eq_of_state} \\
%
  \nabla^2 \phi &=& 4 \pi G \rho_{\rm total}
%                   + 3 p_{total}/c^2) - \Lambda. 
  \label{eq:potential}
\end{eqnarray}
%
The equation of state is shown here for an ideal gas with a ratio of
specific heats $\gamma$.  Equation (\ref{eq:potential}) includes
contributions to the gravitational potential $\phi$ from the total
mass density $\rho_{\rm total}$, which is defined as $\rho_{\rm gas} +
\rho_{\rm dm} + \rho_{\rm stars} - \rho_{\rm average}$.  Although we
write the equations including both MHD and comoving coordinates, the
code is frequently used both in the pure hydrodynamic limit ($\vecB =
0$) and without cosmological expansion ($a = 1$, $\dot{a} = 0$).

@@

\begin{eqnarray}
\vecx & = & \vecr/a, \\
\vecv & = & a \hspace{0.5mm} d\vecx/dt = 
              \vecvp - \dot{a}\vecx, \\
\rho  & = & a^3 \rhop,   \\
p     & = & a^3 \pp, \\
E     & = & a^3 \left(\Ep - 
            \dot{a} \vecx \cdot \vecvp - 
            \frac{1}{2} \dot{a}^2 \vecx^2\right), \\
\phi  & = & \phip + \frac{1}{2} a \ddot{a} \vecx^2. \\
\vecB & = & a^2 \vecB^\prime,
\end{eqnarray}
where primes indicate the usual quantities defined in a fixed frame
frame, $\vecr$ is the proper distance and $\vecx$ is the comoving
distance.

%These definitions result in the following comoving equivalents to eqs. 
%(\ref{eq:mass})-(\ref{eq:potential}).
%%
%
%\begin{eqnarray} 
%\frac{\partial \rhop}{\partial t} 
%          + \frac{1}{a} \vecvp \cdot \nabla \rhop 
%    & = & - \frac{1}{a} \rhop \nabla \cdot \vecvp, 
%        \label{eq:mass_prime} \\
%%
%\frac{\partial \vecvp}{\partial t}  
%          + \frac{1}{a} ( \vecvp \cdot \nabla ) \vecvp 
%    & = & - \frac{\dot{a}}{a} \vecvp 
%          - \frac{1}{a \rhop} \nabla \pp 
%          - \frac{1}{a} \nabla \phip, 
%        \label{eq:momentum_prime} \\
%%
%\frac{\partial \Ep}{\partial t} 
%         + \frac{1}{a} \vecvp \cdot \nabla \Ep 
%   & = & - \frac{\dot{a}}{a} (3 \frac{\pp}{\rhop} + {\vecvp \cdot \vecvp} )
%         - \frac{1}{a \rhop } \nabla \cdot (p \vecvp) 
%%\nonumber   & &   %uncomment for 2-column
%         - \frac{1}{a} \vecvp \cdot \nabla \phip 
%                + \Gamma - \Lambda.
%       \label{eq:total_energy_prime} \\
%%
%\Ep  & = & \ep + \frac{1}{2} {\vecvp \cdot \vecvp},
%        \label{eq:total_energy_def_prime} \\
%%
%\ep                  & = & 
%        \pp/\left[ \left( \gamma - 1 \right) \rhop \right],
%      \label{eq:state_prime} \\ 
%{\nabla}^2 \phip    & = & 
%        \frac{4\pi G}{a} ( \rhop_{total} - \rhop_0 ).
%      \label{eq:poisson}
%\end{eqnarray}
%%
%%We use the proper peculiar baryonic velocity $\vecv_b \equiv a(t) d\vecx / dt$, proper pressure $p$, and modified gravitational potential $ \phi $ which is related to the potential in proper coordinates $\Phi$ by 
%%$$\phi \equiv \Phi + \frac{1}{2} a \ddot{a} \vecx^2. $$
%%The density, however, is comoving:
%%$$\rho_b \equiv \rho_{b,proper}a(t)^3, $$
The expansion parameter $a \equiv 1/(1 + z)$ follows the expansion of
a smooth, homogeneous background, where $z$, the redshift, is a
function only of $t$.  All derivatives are determined with respect to
the comoving position $\vecx$, which is defined simply to remove the
universal expansion from the coordinate system.

The evolution of $a(t)$ is governed by the formula for the expansion
of an isotropic, homogeneous universe:
%
\begin{eqnarray}
\frac{\ddot{a}}{a} & = & 
      - \frac{4 \pi G }{3 a^3 } (\rho_0 
      + 3p_0/c^2) 
      + \Lambda_c /3 .
      \label{eq:expansion} 
\end{eqnarray}
%
Here, $\rho_0$ is the mean comoving mass density (including both
baryonic and dark matter), $p_0$ is the comoving background pressure
contribution, and $\Lambda_c$ is the cosmological constant.
%$\gamma$ is the ratio of specific heats, and $\Lambda$ is the cosmological constant.
%We note that if $\gamma$ is not $5/3$ then $E$ in the first term on
%the right side of eq.~(ref{eq:total_energy}) must be replaced with 
%eq.~(\ref{eq:state}).
This system of equations is limited to the non-relativistic regime and
assumes that curvature effects are not important --- both assumptions
are reasonable as long as the size of the simulated region is small
compared to the radius of curvature and the Hubble length $c/H$, where
$c$ is the speed of light and $H = \dot{a}/a$ is the Hubble constant.

It is clear that the evolution equations are equivalent to the fixed
coordinate version in the limit in which $a = 1$ and $\dot{a} = 0$.
We include the cosmological terms with the understanding that \enzo\
is not restricted to cosmological applications -- while historically
the code was written with cosmological situations in mind (e.g.,
galaxy clusters), in more recent years it has been used to simulate a
much broader range of astrophysical environments.

@@

Any collisionless component (e.g., dark matter, stars) is modeled by
particles which are governed by Newton's equations in comoving
coordinates:
%
\begin{eqnarray}
\frac{d \vecx}{dt} 
    & = & \frac{1}{a} \vecv, 
          \label{eq:dm_position} \\
\frac{d \vecv}{dt} 
    & = & - \frac{\dot{a}}{a} \vecv
          - \frac{1}{a} \nabla \phi, 
          \label{eq:dm_velocity} 
\end{eqnarray}
%
The particles also make a contribution to the gravitational potential
through their contribution to Poisson's equation (Equation
\ref{eq:potential}).

@@

In addition, we can solve the mass conservation equations for a set of
chemical species and reactions.  For any species $i$ with comoving
number density $n_i$, these equations have the form:
\begin{equation}
  \frac{\partial n_i}{\partial t} 
  + \frac{1}{a} \nabla \cdot (n_i \vecv) = 
  k_{ij}(T) n_i n_j + \Gamma^{\rm ph}_j n_j 
  \label{eq:species_evolution}
\end{equation}
where $k_{ij}$ are the rate coefficients for the two-body reactions
and are usually functions of only the gas species (we will
specifically note the cases where we either include three-body
reactions or have density-dependent rates).  The $\Gamma^{\rm ph}_j$
are destruction/creation rates due to photoionizations and/or
photodissociations.  Currently the species we can follow include H,
H$^+$, He, He$^+$, He$^{++}$, and e$^-$, as well as additional options
that add H$^-$, H$_2$, H$_2^+$, and HD, D, and D$^+$.  We also can
track the advection of a comoving metal density that may contribute to
the radiative cooling.

@@

%\begin{equation}
%\frac{\partial \rho_b}{\partial t} + \frac{1}{a} \vecv_b \cdot \nabla \rho_b =  \label{enzoconserve}
% - \frac{1}{a} \rho_b \nabla \cdot \vecv_b 
%\end{equation}

%\begin{equation}
%\frac{\partial \vecv_b}{\partial t} + \frac{1}{a} (\vecv_b \cdot \nabla) \vecv_b = \label{enzomomentum} 
%-\frac{\dot{a}}{a} \vecv_b - \frac{1}{a \rho_b} \nabla p - \frac{1}{a} \nabla \phi
%\end{equation}

%\begin{eqnarray}
%\frac{\partial E}{\partial t} + \frac{1}{a} \vecv_b \cdot \nabla E =  
%- \frac{\dot{a}}{a} \left( 3 \frac{p}{\rho_b} + \vecv_b^2 \right)  \nonumber \\
%- \frac{1}{a \rho_b} \nabla \cdot (p \vecv_b) \nonumber \\
%- \frac{1}{a} \vecv_b \cdot \nabla \phi + \Gamma - \Lambda
%\label{enzoenergy}
%\end{eqnarray}

@@

%\begin{equation}
%\frac{d \vecv_{part}}{dt} = -\frac{\dot{a}}{a} \vecv_{part} - \frac{1}{a} \nabla \phi
%\label{enzopartmom}
%\end{equation}

@@

The code can include either a homogeneous radiation background (with
possible local self-shielding corrections), or evolve an inhomogeneous
radiation field either by directly solving the radiative transfer
equation along rays or by solving a set of moment equations derived
from the radiative transfer equation.

The radiative transfer equation in comoving coordinates
\citep[e.g.,][]{Gnedin97} reads
%
\begin{equation}
  \label{eq:rteqn}
  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
  \frac{\hat{n} \cdot \nabla I_\nu}{\bar{a}} -
  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
  3 I_\nu \right) = -\kappa_\nu I_\nu + j_\nu .
\end{equation}
%
Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is the radiation
specific intensity in units of energy per time $t$ per solid angle per
unit area per frequency $\nu$.  $\bar{a} = a/a_{em}$ is the ratio of
scale factors at the current time and time of emission.  The second
term represents the propagation of radiation, where the factor
$1/\bar{a}$ accounts for cosmic expansion.  The third term describes
both the cosmological redshift and dilution of radiation.  On the
right hand side, the first term considers the absorption coefficient
$\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$, and the second term
$j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ is the emission coefficient
that includes any sources of radiation (whether point or diffuse).

@@

As an alternative to the ray-casting strategy, we solve an equation
derived from integrating the equation of radiative transfer over
angles.  In particular, we use a flux-limited diffusion approximation
for radiative transfer, with couplings to both the gas energy and
chemical number densities. The equations solved are:
\begin{eqnarray}
  \label{eq:fld_radiation}
  \frac{\partial E_r}{\partial t} + \frac1a \nabla\cdot\left(E_r \vecv \right) &=& 
  \nabla\cdot\left(D\nabla E_r\right) -
  \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
  \label{eq:fld_heating}
  \frac{\partial e_c}{\partial t} &=& -\frac{2\dot{a}}{a} e_c + \Gamma - \Lambda,
%  \label{eq:fld_chemistry}
%  \partial_t {\mathrm n}_i + 
%    \frac1a \nabla\cdot\left({\mathrm n}_i \vecv_b\right) &=& 
%    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
%    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
\end{eqnarray}
where $E_r = E_r(\vecx,t)$ is a grey radiation energy density and
$e_c$ is the internal energy correction due to photo-heating and
chemical cooling.
%and ${\mathrm n}_i$ correspond to the chemical number densities of
%HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
%number density.  
Here, we define the grey radiation energy density through first
assuming a fixed frequency spectrum, i.e.  $E_{\nu}(\nu,\vecx,t) =
\tilde{E}_r(\vecx,t) \chi(\nu)$, and then defining the integrated
quantity
\begin{equation}
\label{eq:grey_radiation_energy}
   E_r(\vecx,t) \equiv \int_{\nu_0}^{\infty}
   E_{\nu}(\nu,\vecx,t)\,\mathrm d\nu \  = \ 
   \tilde{E}_r(\vecx,t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
\end{equation}
In addition, $D$ in Equation (\ref{eq:fld_radiation}) is the Larsen
square-root flux-limiter \citep[see][]{Morel2000}; $\kappa$ is the
total opacity; $\eta$ is the field of radiation sources; $\Gamma$
provides the radiation induced photo-heating, and $\Lambda$ the
chemical cooling.
%, $\alpha_{i,j}$ are the reaction rate coefficients defining
%the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
%and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.

@@

In this section, we briefly describe the numerical methods which are
used to solve the equations outlined in Section~\ref{sec.equations}.
We proceed through the numerical methods in the same order as will be
used in Sections~\ref{sec.amr} through~\ref{sec.num.analysis} so that
there is a one-to-one correspondence between each of the following
overviews and the complete description provided in the later sections.
The goal in this section is to introduce the reader to the basic
principles of the methods without drowning in detail.

>>>>

We begin this section by first writing down the complete set of
physical equations that can be solved by \enzo, and then briefly
describe the numerical algorithms that we use to solve these
equations.  This section is intended to be an overview of
\enzo's capabilities: thus, we gather all of the equations
solved into a single place and provide a brief and high-level
introduction to the numerics of the code.  Detailed descriptions of
the individual components are then provided in
Sections~\ref{sec.amr}-\ref{sec.num.analysis}. In
Table~\ref{tab:variables} we provide a convenient summary of all
variables and physical constants used throughout this manuscript.

@@

The Eulerian equations of ideal magnetohydrodynamics (MHD) including
gravity, in a coordinate systems comoving with the cosmological
expansion, are given by

\begin{eqnarray} 
  \frac{\partial \rho}{\partial t} 
  + \frac{1}{a} \div (\rho \vecv) & = & 0,
  \label{eq:mass} \\
%
  \frac{\partial \rho \vecv}{\partial t}  
  + \frac{1}{a} \div \left(\rho \vecv \vecv + \myvec{I}p^* - 
    \frac{\vecB \vecB}{a} \right) & = &
  - \frac{\dot{a}}{a} \rho \vecv - \frac{1}{a} \rho \grad \phi,
% 
%          + \frac{1}{a} ( \vecv \cdot \grad ) \vecv
%    & = & - \frac{\dot{a}}{a} \vecv 
%          - \frac{1}{a \rho} \grad p
%          - \frac{1}{a \rho} (\nalba \times \vecB) \times \vecB
%          - \frac{1}{a} \grad \phi, 
  \label{eq:momentum} \\
%
  \frac{\partial E}{\partial t} 
  + \frac{1}{a} \div \left[ (E + p^*) \vecv - 
    \frac{1}{a} \vecB (\vecB \cdot \vecv) \right] & = &
  - \frac{\dot{a}}{a} \left( 2E - \frac{B^2}{2a} \right) - 
  \frac{1}{a} \vecv \cdot \grad \phi 
  - \Lambda + \Gamma + \frac{1}{a^2} \div \fcond,
% 
%         + \frac{1}{a} \vecv \cdot \grad E
%  & = & - \frac{\dot{a}}{a} (3 \frac{p}{\rho} + {\vecv \cdot \vecv} )
%       - \frac{1}{a \rho } \div (p \vecv) 
%%\nonumber   & &   %uncomment for 2-column
%        - \frac{1}{a} \vecv \cdot \grad \phi 
%               + \Gamma - \Lambda.
  \label{eq:total_energy}  \\
%
  \frac{\partial \vecB}{\partial t} - 
  \frac{1}{a}  \grad \times (\vecv \times \vecB) & = & 
  -\frac{\dot{a}}{2 a} \vecB. \label{eq:induction}
\end{eqnarray}

%\begin{eqnarray} 
%\frac{\partial \rho}{\partial t} 
%          + \vecv \cdot \grad \rho
%    & = & - \rho \div \vecv,
%        \label{eq:mass} \\
%%
%\frac{\partial \vecv }{ \partial t} 
%          + ( \vecv \cdot \grad ) \vecv 
%    & = & - \frac{1}{\rho} \grad p
%          - \grad \phi,
%        \label{eq:momentum} \\
%%
% \frac{\partial E}{\partial t} + \vecv \cdot \grad E
%    & = & - \frac{1}{\rho} \div (p \vecv) 
%          - \vecv \cdot \grad \phi
%%\nonumber          \\  & &     % uncomment for 2-column
%     + \Gamma - \Lambda
%        \label{eq:total_energy}
%\end{eqnarray}
%
In these equations, $E$, $\rho$, $\vecv$, and $\vecB$ are the comoving
total fluid energy density, comoving gas density, peculiar
velocity, and comoving magnetic field strength, respectively.
$\myvec{I}$ is the identity matrix, and $a$ is the cosmological
expansion parameter (discussed in more detail below).  The first
equation represents conservation of mass, the second conservation of
momentum, and the third conservation of total (kinetic plus thermal)
fluid energy.  They are respectively, the first, second, and third
moments of the Boltzmann equation.  The fourth equation is the
magnetic induction equation.  Terms representing radiative cooling
($\Lambda$) and heating ($\Gamma$) enter on the right-hand side of the
energy equation (\ref{eq:total_energy}), as does the flux due to
thermal heat conduction ($\fcond$).

The total fluid energy density $E$ is given by
\begin{equation}
E =  e + \frac{\rho v^2}{2}  + \frac{B^2}{2a}
        \label{eq:total_energy_def},
\end{equation}
where $e$ is the comoving thermal energy density, the total
isotropic pressure $p^*$ is given by
\begin{equation}
p^* = p + \frac{B^2}{2a},
\end{equation}
and $p$ is the thermal pressure.  The equations are closed by an
equation of state and Poisson's equation (in the Newtonian limit) for
the gravitational potential $\phi$:
%
\begin{eqnarray}
%
  e &=& \frac{p}{(\gamma - 1) \rho},
  \label{eq:eq_of_state} \\
%
  \nabla^2 \phi &=& 4 \pi G \left( \rho_{\rm total} - \rho_0 \right)
%                   + 3 p_{total}/c^2) - \Lambda. 
  \label{eq:potential}
\end{eqnarray}
%
The equation of state is shown here for an ideal gas with a ratio of
specific heats $\gamma$.  The gravitational potential $\phi$ is
sourced by the total mass density contrast, where $\rho_{\rm total} =
\rho_{\rm gas} + \rho_{\rm dm} + \rho_{\rm stars}$ and $\rho_0$ is the
mean density.  Although we write the equations including both magnetic
field terms and comoving coordinates, the code is frequently used both
in the purely hydrodynamic limit ($\vecB = 0$, referred to as HD
below) and without cosmological expansion ($a = 1$, $\dot{a} = 0$).

@@

\begin{eqnarray}
\vecx & = & \vecx' / a, \\
\vecv & = & a \hspace{0.5mm} \dot{\vecx} = 
              \vecvp - \dot{a}\vecx, \\
\rho  & = & a^3 \rhop,   \\
p     & = & a^3 \pp, \\
E     & = & a^3 \left(\Ep - 
            \dot{a} \, \vecx \cdot \vecvp - 
            \frac{1}{2} \dot{a}^2 \vecx^2\right), \\
\phi  & = & \phip + \frac{1}{2} \, a \ddot{a} \, \vecx^2, \\
\vecB & = & a^2 \vecB^\prime,
\end{eqnarray}
where primes indicate quantities in proper coordinates (fixed frame).
%These definitions result in the following comoving equivalents to eqs. 
%(\ref{eq:mass})-(\ref{eq:potential}).
%%
%
%\begin{eqnarray} 
%\frac{\partial \rhop}{\partial t} 
%          + \frac{1}{a} \vecvp \cdot \grad \rhop 
%    & = & - \frac{1}{a} \rhop \div \vecvp, 
%        \label{eq:mass_prime} \\
%%
%\frac{\partial \vecvp}{\partial t}  
%          + \frac{1}{a} ( \vecvp \cdot \grad ) \vecvp 
%    & = & - \frac{\dot{a}}{a} \vecvp 
%          - \frac{1}{a \rhop} \grad \pp 
%          - \frac{1}{a} \grad \phip, 
%        \label{eq:momentum_prime} \\
%%
%\frac{\partial \Ep}{\partial t} 
%         + \frac{1}{a} \vecvp \cdot \grad \Ep 
%   & = & - \frac{\dot{a}}{a} (3 \frac{\pp}{\rhop} + {\vecvp \cdot \vecvp} )
%         - \frac{1}{a \rhop } \div (p \vecvp) 
%%\nonumber   & &   %uncomment for 2-column
%         - \frac{1}{a} \vecvp \cdot \grad \phip 
%                + \Gamma - \Lambda.
%       \label{eq:total_energy_prime} \\
%%
%\Ep  & = & \ep + \frac{1}{2} {\vecvp \cdot \vecvp},
%        \label{eq:total_energy_def_prime} \\
%%
%\ep                  & = & 
%        \pp/\left[ \left( \gamma - 1 \right) \rhop \right],
%      \label{eq:state_prime} \\ 
%{\nabla}^2 \phip    & = & 
%        \frac{4\pi G}{a} ( \rhop_{total} - \rhop_0 ).
%      \label{eq:poisson}
%\end{eqnarray}
%%
%%We use the proper peculiar baryonic velocity $\vecv_b \equiv a(t) d\vecx / dt$, proper pressure $p$, and modified gravitational potential $ \phi $ which is related to the potential in proper coordinates $\Phi$ by 
%%$$\phi \equiv \Phi + \frac{1}{2} a \ddot{a} \vecx^2. $$
%%The density, however, is comoving:
%%$$\rho_b \equiv \rho_{b,proper}a(t)^3, $$
The expansion parameter $a \equiv 1/(1 + z)$ follows the expansion of
a smooth, homogeneous background, where $z$, the redshift, is a
function only of $t$.  All spatial derivatives are determined with
respect to the comoving position $\vecx$, which removes the universal
expansion from the coordinate system. The evolution of $a(t)$ is
governed by the second Friedmann equation for the expansion of a
spatially homogeneous and isotropic universe
%
\begin{eqnarray}
\frac{\ddot{a}}{a} & = & 
      - \frac{4 \pi G }{3 a^3 } (\rho_0 
      + 3p_0/c^2) 
      + \Lambda_c c^2 / 3 .
      \label{eq:expansion} 
\end{eqnarray}
%
Here $\rho_0$ is the mean comoving mass density (including both
baryonic and dark matter), $p_0$ is the comoving background pressure
contribution, and $\Lambda_c$ is the cosmological constant.
%$\gamma$ is the ratio of specific heats, and $\Lambda$ is the cosmological constant.
%We note that if $\gamma$ is not $5/3$ then $E$ in the first term on
%the right side of eq.~(ref{eq:total_energy}) must be replaced with 
%eq.~(\ref{eq:state}).
This system of equations is limited to the non-relativistic regime and
assumes that curvature effects are not important --- both assumptions
are reasonable as long as the size of the simulated region is small
compared to the radius of curvature and the Hubble length $c/H$, where
$c$ is the speed of light and $H = \dot{a}/a$ is the Hubble constant.

The comoving evolution equations are equivalent to the fixed
coordinate version in the non-cosmological limit ($a = 1$, $\dot{a} =
0$).  We include the cosmological terms with the understanding that
\enzo\ is not restricted to cosmological applications -- while
historically the code was written with cosmological situations in mind
(e.g. galaxy clusters), in more recent years it has been used to
simulate a much broader range of astrophysical environments.

@@

Any collisionless components (e.g. dark matter and stars) are modeled
by N-body particles, whose dynamics are governed by Newton's equations
in comoving coordinates:
%
\begin{eqnarray}
\frac{d \vecx}{dt} 
    & = & \frac{1}{a} \vecv, 
          \label{eq:dm_position} \\
\frac{d \vecv}{dt} 
    & = & - \frac{\dot{a}}{a} \vecv
          - \frac{1}{a} \grad \phi, 
          \label{eq:dm_velocity} 
\end{eqnarray}
%
The particles contribute to the gravitational potential through
Poisson's equation (Equation \ref{eq:potential}).

@@

In addition, \enzo\ can solve the mass conservation equations for a set of
chemical species and their reactions.  For any species $i$ with comoving
number density $n_i$, these equations have the form:
\begin{equation}
  \frac{\partial n_i}{\partial t} 
  + \frac{1}{a} \div (n_i \vecv) = 
  \sum_j k_{ij}(T) \, n_i n_j + \sum_j \Gamma^{\rm ph}_j n_j 
  \label{eq:species_evolution}
\end{equation}
where $k_{ij}$ are the rate coefficients for the two-body reactions
and are usually functions of only the gas species (we will
specifically note the cases where we either include three-body
reactions or have density-dependent rates).  The $\Gamma^{\rm ph}_j$
are destruction/creation rates due to photoionizations and/or
photodissociations.  Currently the species \enzo\ can follow include
H, H$^+$, He, He$^+$, He$^{++}$, and e$^-$, and through additional
options also H$^-$, H$_2$, H$_2^+$, and HD, D, and D$^+$. Lastly,
\enzo\ can also track the advection of a comoving metal density field,
which can contribute to the radiative cooling and star formation
processes.

@@

%\begin{equation}
%\frac{\partial \rho_b}{\partial t} + \frac{1}{a} \vecv_b \cdot \grad \rho_b =  \label{enzoconserve}
% - \frac{1}{a} \rho_b \div \vecv_b 
%\end{equation}

%\begin{equation}
%\frac{\partial \vecv_b}{\partial t} + \frac{1}{a} (\vecv_b \cdot \grad) \vecv_b = \label{enzomomentum} 
%-\frac{\dot{a}}{a} \vecv_b - \frac{1}{a \rho_b} \grad p - \frac{1}{a} \grad \phi
%\end{equation}

%\begin{eqnarray}
%\frac{\partial E}{\partial t} + \frac{1}{a} \vecv_b \cdot \grad E =  
%- \frac{\dot{a}}{a} \left( 3 \frac{p}{\rho_b} + \vecv_b^2 \right)  \nonumber \\
%- \frac{1}{a \rho_b} \div (p \vecv_b) \nonumber \\
%- \frac{1}{a} \vecv_b \cdot \grad \phi + \Gamma - \Lambda
%\label{enzoenergy}
%\end{eqnarray}

@@

%\begin{equation}
%\frac{d \vecv_{part}}{dt} = -\frac{\dot{a}}{a} \vecv_{part} - \frac{1}{a} \grad \phi
%\label{enzopartmom}
%\end{equation}

@@

The code can include either a homogeneous radiation background (with
possible local self-shielding corrections), or evolve an inhomogeneous
radiation field either by directly solving the radiative transfer
equation along rays or by solving a set of moment equations derived
from the radiative transfer equation. In comoving coordinates
\citep[e.g.,][]{Gnedin97} the radiative transfer equation reads
%
\begin{equation}
  \label{eq:rteqn}
  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
  \frac{a_{\rm em}}{a} \, \hat{n} \cdot \grad I_\nu -
  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
  3 I_\nu \right) = -\kappa_\nu I_\nu + j_\nu .
\end{equation}
%
Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is specific
intensity of the radiation, with dimensions of energy per time $t$ per
solid angle $\Omega$ per unit area per frequency $\nu$.  The second
term represents the propagation of radiation, where the factor $a_{\rm
  em}/a$ accounts for cosmic expansion since the time of emission.
The third term describes both the cosmological redshift and dilution
of radiation.  On the right hand side, the first term captures
absorption ($\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$), and the
second term emission ($j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ from any
sources of radiation (whether point or diffuse).

@@

As an alternative to the ray-casting strategy, \enzo\ can also solve
the angle-averaged radiative transfer equation, through a
flux-limited diffusion approximation with couplings to both the gas
energy and chemical number densities. The equations solved are
\begin{eqnarray}
  \label{eq:fld_radiation}
  \frac{\partial E_r}{\partial t} + \frac1a \div\left(E_r \vecv \right) &=& 
  \div\left(D\grad E_r\right) -
  \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
  \label{eq:fld_heating}
  \frac{\partial e_c}{\partial t} &=& -\frac{2\dot{a}}{a} e_c + \Gamma - \Lambda,
%  \label{eq:fld_chemistry}
%  \partial_t {\mathrm n}_i + 
%    \frac1a \div\left({\mathrm n}_i \vecv_b\right) &=& 
%    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
%    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
\end{eqnarray}
where $E_r = E_r(\vecx,t)$ is a grey radiation energy density and
$e_c$ is the internal energy correction due to photo-heating and
chemical cooling.
%and ${\mathrm n}_i$ correspond to the chemical number densities of
%HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
%number density.  
Here, we define $E_r$ through first assuming a fixed frequency
spectrum, i.e.  $E_{\nu}(\nu,\vecx,t) = \tilde{E}_r(\vecx,t)
\chi(\nu)$, and then defining the integrated quantity
\begin{equation}
\label{eq:grey_radiation_energy}
   E_r(\vecx,t) \equiv \int_{\nu_0}^{\infty}
   E_{\nu}(\nu,\vecx,t)\,\mathrm d\nu \  = \ 
   \tilde{E}_r(\vecx,t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
\end{equation}
$D$ in Equation (\ref{eq:fld_radiation}) is the Larsen
square-root flux-limiter \citep[see][]{Morel2000}, $\kappa$ the
total opacity, $\eta$ the field of radiation sources, $\Gamma$
the radiation induced photo-heating rate, and $\Lambda$ the
chemical cooling rate.
%, $\alpha_{i,j}$ are the reaction rate coefficients defining
%the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
%and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.

@@

In this section we briefly describe the numerical methods which are
used to solve the equations outlined in Section~\ref{sec.equations}.
We proceed through the numerical methods in the same order as will be
used in Sections~\ref{sec.amr} through~\ref{sec.num.analysis} so that
there is a one-to-one correspondence between each of the following
overviews and the complete description provided in the later sections.
The goal in this section is to introduce the reader to the basic
principles of the methods without drowning in detail.
